### 2018-04-17 paper.Rmd

Gender differences in language production have been found in multiple studies, notably in @eriksson2012differences, a large-scale study that spanned ten languages. <!-- SM what research questions specifically? many aspects of gender in language... probably better to note the range of gender-related effects in early language -->
A researcher or student in child language acquisition may now investigate this difference with relative ease using `childesr`. <!-- SM: unclear what "this difference" refers to -->
By pulling data from the `transcript_by_speaker` table, a user has access to the the multitude of derived linguistic measures that are often used to evaluate a child's linguistic performance. <!-- SM: linguistic performance -> grammatical development? -->
In this worked example we walk through a sample analysis that probes the gender differences in language production. <!-- SM: language production > gender differences in early grammatical development -->

First, we utilize the `childesr` function call `get_speaker_statistics` to pull data relating to the aforementioned derived measures for children and their transcripts.
Notee that we exclusively select the data of children, and exclude their caregivers.

@@

This call pulls data from all collections and corpora, including those languages for which there are very sparse data.
In order to make any substantial inferences from our analysis, we begin by filterering languages for which there are a significant number of children. 
We also filter for children under the age of four years.

@@

The data contained in CHILDES is populated from a diverse array of studies that adhered to varying conditions and paradigms.
This point is particularly salient in our gender analysis due to potential non-independence issues that result from longitudinal studies.
We chose to fit a model (`mod_data`) of this data that takes these facts into account, and for brevity, is not included here.^[All code and analyses are available at https://github.com/langcog/childes-db-paper] The resultant plot displays the average MTLD scores for various children at different ages, split by gender, with a line corresponding to our model.

@@

This plot reveals a slight gender difference in linguistic productivity in young children, replicating the results of @eriksson2012differences.
The goal of this analysis was to showcase an example of the the steps involved in using `childesr` for exploring the CHILDES dataset, as well as some of the potential hazards in working with a diverse set of corpora.

<!-- SM: not terribly clear what the hazards were: sparsity and autocorrelation? -->

>>>>

Gender differences in language production have been found in multiple studies, notably in @eriksson2012differences, a large-scale study that spanned ten languages and found that females have a higher vocabulary than males in early childhood.
A researcher or student in child language acquisition may now investigate this difference with relative ease using `childesr`.
By pulling data from the `transcript_by_speaker` table, a user has access to the the multitude of derived linguistic measures that are often used to evaluate a child's grammatical development.
In this worked example we walk through a sample analysis that probes the gender differences in early grammatical development.

First, we utilize the `childesr` function call `get_speaker_statistics` to pull data relating to the aforementioned derived measures for children and their transcripts.
Note that we exclusively select the data of children, and exclude their caregivers.

@@

This call pulls data from all collections and corpora, including those languages for which there are very sparse data.
In order to make any substantial inferences from our analysis, we begin by filtering languages for which there are a significant number of children. 
We also filter for children under the age of four years.

@@

The data contained in CHILDES is populated from a diverse array of studies that adhered to varying conditions and paradigms.
This point is particularly salient in our gender analysis due to potential non-independence issues that result from longitudinal studies.
For this reason, we fit a linear fixed effects model (`mod_data`, shown in the plot below) to remedy these concerns.^[All code and analyses are available at https://github.com/langcog/childes-db-paper] 
The plot below displays the average MTLD scores for various children at different ages, split by gender, with a line corresponding to our mixed effects model.

@@

This plot reveals a slight gender difference in linguistic productivity in young children (displaying a moderate female advantage), replicating the results of @eriksson2012differences.
The goal of this analysis was to showcase an example of the the steps involved in using `childesr` for exploring the CHILDES dataset, as well as some of the potential hazards in working with a diverse set of corpora, many of which are longitudinal.

<!-- SM: not terribly clear what the hazards were: sparsity and autocorrelation? AS: mostly the fact that there are longitudinal studies in here-->
